### 2015-02-07 cav-2015-tex/appendix.tex
Fixed some multiply-defined references
edits

\section*{Appendix: Common Rules for Javalite}
\input{javalite-common}

The surface syntax and the machine syntax for the Javalite syntactic
machine are in the main paper. \figref{fig:javalite-common} defines
the rewrite relations for the portion of the Javalite language
semantics that are common to both generalized symbolic execution and
the summary heap algorithm. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.

@@

\section{Summary Heap Algorithm}

\begin{figure*}[t]
\begin{center}
\setlength{\tabcolsep}{60pt}
\hspace*{-35pt}
\begin{tabular}[c]{cc}
\scalebox{1.0}{\usebox{\boxPFAFW}} & 
%\scalebox{0.91}{\input{faYHeap.pdf_t}} &
\scalebox{0.91}{\input{fwXHeap.pdf_t}} \\ \\
(a) & (b)
\end{tabular}
\end{center}
\caption{Field read and write relations with an example heap. (a) Field-access, $\rsym^\mathit{A}$, and field-write, $\rsym^\mathit{W}$, rewrite rules for the $\rsym$ relation. (b) The final heap after $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\ \cfgt{:=}\ \lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp\rp$.}
\label{fig:fHeap}
\end{figure*}

\begin{figure*}
\begin{tabular}[c]{l}
\scalebox{1.0}{\usebox{\boxPI}} \\
\end{tabular}
\caption{The summary machine, $s ::= \lp\cfgnt{L}\ \cfgnt{R}\ \cfgnt{r}\ \cfgnt{f}\ \cfgnt{C}\rp$, with $s\rsum^*s^\prime =  s \rsum \cdots \rsum s^\prime \rsum s^\prime$.}
\label{fig:symInit}
\end{figure*}

The rewrite rules for field access or field write when a null reference is possible leads to an error state.
$$
\begin{array}{c}
	\inferrule[Field Access (NULL)]{
      \exists \lp \phi\ l\rp \in \cfgnt{L}\lp \cfgnt{r}\rp\ \lp l = l_{\mathit{null}} \wedge \mathbb{S}\lp \phi \wedge \phi_g\rp \rp
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgnt{r}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp  \rsym^\mathit{A^\prime}
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgt{error}\ \cfgt{end} \rp
	} \\
\\
	\inferrule[Field Write (NULL)]{
      \cfgnt{r}_x = \eta\lp \cfgnt{x}\rp \\
      \exists \lp \phi\ l\rp \in \cfgnt{L}\lp \cfgnt{r}_x\rp\ \lp l \neq l_{\mathit{null}} \wedge \mathbb{S}\lp \phi \wedge \phi_g\rp \rp
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgnt{r}\ \lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgt{*}\ \rightarrow\ \cfgnt{k}\rp \rp  \rsym^\mathit{F^\prime}
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgt{error}\ \cfgt{end}\rp
	}	
\end{array}
$$
The rules reflect a branch in the control flow for the summary heap
algorithm: one branch being the feasible null outcome (shown here), and the
other branch being the potential non-null
outcome (shown in the main paper). In the non-null case, the path constraint is updated to
restrict out all possible null instances.

\input{definitions}

>>>>

\section{Rules for Javalite}\label{app:javalite}

\input{javalite-common}

The surface syntax and the machine syntax for the Javalite syntactic
machine are in the main paper. \figref{fig:javalite-common} defines
the rewrite relations for the portion of the Javalite language
semantics that are common to both generalized symbolic execution and
the symbolic heap algorithm. The relation $\rcom$ includes all of these
rules. Excepting \textrm{N{\footnotesize EW}}, the rules do not update
the heap, and are largely concerned with argument evaluation in an
expected way. It is assumed that only type safe programs are input to
the machine so there is no type checking or error conditions. The
machine simply halts if no rewrite is enabled.

@@

\section{Symbolic Heap Algorithm}

%\begin{figure*}[t]
%\begin{center}
%\setlength{\tabcolsep}{60pt}
%\hspace*{-35pt}
%\begin{tabular}[c]{cc}
%\scalebox{1.0}{\usebox{\boxPFAFW}} & 
%%\scalebox{0.91}{\input{faYHeap.pdf_t}} &
%\scalebox{0.91}{\input{fwXHeap.pdf_t}} \\ \\
%(a) & (b)
%\end{tabular}
%\end{center}
%\caption{Field read and write relations with an example heap. (a) Field-access, $\rsym^\mathit{A}$, and field-write, $\rsym^\mathit{W}$, rewrite rules for the $\rsym$ relation. (b) The final heap after $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\ \cfgt{:=}\ \lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp\rp$.}
%\label{fig:fHeap}
%\end{figure*}

%\begin{figure*}
%\begin{tabular}[c]{l}
%\scalebox{1.0}{\usebox{\boxPI}} \\
%\end{tabular}
%\caption{The summary machine, $s ::= \lp\cfgnt{L}\ \cfgnt{R}\ \cfgnt{r}\ \cfgnt{f}\ \cfgnt{C}\rp$, with $s\rsum^*s^\prime =  s \rsum \cdots \rsum s^\prime \rsum s^\prime$.}
%\label{fig:symInit}
%\end{figure*}

The rewrite rules for field access or field write when a null reference is possible leads to an error state.
$$
\begin{array}{c}
	\inferrule[Field Access (NULL)]{
      \exists \lp \phi\ l\rp \in \cfgnt{L}\lp \cfgnt{r}\rp\ \lp l = l_{\mathit{null}} \wedge \mathbb{S}\lp \phi \wedge \phi_g\rp \rp
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgnt{r}\ \lp \cfgt{*}\ \cfgt{\$}\ \cfgnt{f} \rightarrow \cfgnt{k}\rp \rp  \rsym^\mathit{A^\prime}
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgt{error}\ \cfgt{end} \rp
	} \\
\\
	\inferrule[Field Write (NULL)]{
      \cfgnt{r}_x = \eta\lp \cfgnt{x}\rp \\
      \exists \lp \phi\ l\rp \in \cfgnt{L}\lp \cfgnt{r}_x\rp\ \lp l \neq l_{\mathit{null}} \wedge \mathbb{S}\lp \phi \wedge \phi_g\rp \rp
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgnt{r}\ \lp \cfgnt{x}\ \cfgt{\$}\ \cfgnt{f}\ \cfgt{:=}\ \cfgt{*}\ \rightarrow\ \cfgnt{k}\rp \rp  \rsym^\mathit{F^\prime}
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi_g\ \eta\ \cfgt{error}\ \cfgt{end}\rp
	}	
\end{array}
$$ The rules reflect a branch in the control flow for the symbolic
heap algorithm: one branch being the feasible null outcome (shown
here), and the other branch being the potential non-null outcome
(shown in the main paper). In the non-null case, the path constraint
is updated to restrict out all possible null instances.

%\input{definitions}

### 2015-02-07 cav-2015-tex/conclusion.tex
Added the technnical stuff
edits

Although our approach is inspired by the work in the static analysis
domain, the heap representations used by those techniques do not lend
themselves to symbolic execution since they rely on
approximations, they are inflexible in what types of relations can be
expressed, and they rely on rewriting of constraints.  In this work,
we present a heap summary technique for symbolic execution that
maintains exactness, represents arbitrary heap structures without
restrictions on aliasing relationships, and supports dynamic
on-the-fly initialization of heap structures. This unique
combination of capabilities enables the summary heap to outperform
state-of-the-art symbolic execution techniques, and in some cases, the
performance gains are considerable. We believe that performance gains
in symbolic execution can naturally benefit a variety of analyses
based on symbolic execution, for example, verification of properties,
especially those related to the heap, test case generation, program
evolution techniques such as directed incremental symbolic
execution~\cite{person:pldi2011}, and other back-end analyses.

>>>>

Although our approach is inspired by the work in the static analysis
domain, the heap representations used by those techniques do not lend
themselves to symbolic execution since they rely on approximations,
they are inflexible in what types of relations can be expressed, and
they rely on rewriting of constraints.  In this work, we present a
heap representation for symbolic execution that maintains exactness,
represents arbitrary heap structures without restrictions on aliasing
relationships, and supports dynamic on-the-fly initialization of heap
structures. This unique combination of capabilities enables the
symbolic heap algorithm to outperform state-of-the-art symbolic
execution techniques, and in some cases, the performance gains are
considerable. We believe that performance gains in symbolic execution
can naturally benefit a variety of analyses based on symbolic
execution, for example, verification of properties, especially those
related to the heap, test case generation, program evolution
techniques such as directed incremental symbolic
execution~\cite{person:pldi2011}, and other back-end analyses. In
future work we plan to evaluate the impact of analyses that rely on
symbolic execution.

### 2015-02-07 cav-2015-tex/evaluation.tex
edits

>>>>

The prototype of our~\symtxt{} algorithm is implemented as a new
extension to SPF.  In addition to the operations presented in the in
this paper, the implementation contains support for operations over
integers.  We are currently working on adding support for floating
point operations, arrays, and bit-operations.  We use the
\texttt{jConstraints} library which provides support for handling heap
constraints~\cite{ase2014-ghilrr,jpf2014-dghirr}. It uses the z3
solver~\cite{deMouraBjorner08Z3} as the underlying constraint
solver. We leverage the incremental solving feature in z3 in our
experiments. We also implement other optimizations that cache and
re-use solutions to previous constraints. These optimizations are an
important part of our implementation because only small portions of
the heap constraint changes from one state to another.

### 2015-02-07 cav-2015-tex/summary-heap.tex
edits

The soundness and completeness proofs in Section~\ref{sec:bisim} rely on a set of 
properties of the symbolic heap that
are preserved by the small-step semantics during symbolic execution:
%\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
%\emph{type consistency}.
\begin{compactdesc}
%<<<<<<< HEAD
\item[Reference partitions:] Each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}% that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in constraints)
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}% that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere)
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} %(these appear in environments, expressions, and continuations).
\item[Null and Uninitialized:] Every summary heap contains 
two special locations, null ($\cfgnt{l}_\mathit{null}$), and uninitialized,
%=======
%\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
%  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
%  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
%  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
%  literals} (these appear in environments, expressions, and continuations).
%\item[Null and Uninitialized] means that every symbolic heap contains 
%a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
%location for uninitialized
%>>>>>>> 49c83f34a0b5afd58adc4b6649b15e2cbc93437c
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability:] References do not change once added to the
location map, eliminating any need for constraint rewriting as the program state evolves.
\item[Determinism:] A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency:] All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

Other more complex rules, such as one to create a new instance of a
class, define constraints on the rewrites and more complex
transformations on the heap.
$$
\mprset{flushleft}
	\inferrule[New]{
      \cfgnt{r} = \mathrm{stack}_r\lp \rp\\
      l = \mathrm{fresh}_l\lp \cfgnt{C}\rp\\\\
      \cfgnt{R}^\prime = \cfgnt{R}[\forall \cfgnt{f} \in \mathit{fields}\lp \mathrm{C}\rp \ \lp \lp l\ \cfgnt{f}\rp  \mapsto \cfgnt{r}_\mathit{null} \rp ] \\\\
      \cfgnt{L}^\prime = \cfgnt{L}[\cfgnt{r} \mapsto \{\lp \cfgt{true}\ l\rp \}]
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgt{new}\ \cfgnt{C}\rp \ \cfgnt{k}\rp  \rcom
      \lp \cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \phi\ \eta\ \cfgnt{r}\ \cfgnt{k}\rp 
	}
$$ In this rule, when the string matches the new-expression, it is
        rewritten to use a new heap location where all of the fields
        for the new object point to $\cfgnt{r}_\mathit{null}$ and the
        location map points a new stack reference to that new object.
        Note that references for objects from the new-expression are
        auxiliary literals, so they do not appear in any constraint in
        the symbolic heap as an aliasing option.

>>>>

The soundness and completeness proofs in Section~\ref{sec:bisim} rely
on certain characteristics of the symbolic heap that are preserved by
the small-step semantics during symbolic execution:
%\emph{reference partitions}, \emph{null and uninitialized}, \emph{immutability}, \emph{determinism}, and
%\emph{type consistency}.
\begin{compactdesc}
%<<<<<<< HEAD
\item[Reference partitions:] Each reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
  heap}% that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in constraints)
  ; $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
  literals}% that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere)
  ; and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
  literals} %(these appear in environments, expressions, and continuations).
\item[Null and Uninitialized:] Every summary heap contains 
two special locations, null ($\cfgnt{l}_\mathit{null}$), and uninitialized,
%=======
%\item[Reference partitions] means that a reference belongs to one of three partitions: $\mathrm{init}_\cfgnt{r}()$ for the \emph{input
%  heap} that includes $\cfgnt{r}_\mathit{null}$ (these are created by lazy initialization and are the only references that appear as variables in
%  constraints); $\mathrm{fresh}_\cfgnt{r}()$ for \emph{auxiliary
%  literals} that includes $\cfgnt{r}_\mathit{un}$ (these add structure but do not appear elsewhere); and $\mathrm{stack}_\cfgnt{r}()$ for \emph{stack
%  literals} (these appear in environments, expressions, and continuations).
%\item[Null and Uninitialized] means that every symbolic heap contains 
%a special location for null ($\cfgnt{l}_\mathit{null}$), and a special
%location for uninitialized
%>>>>>>> 49c83f34a0b5afd58adc4b6649b15e2cbc93437c
  ($\cfgnt{l}_\mathit{un}$) with $\cfgnt{L}(\cfgnt{r}_\mathit{null}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{null})\}$ and
  $\cfgnt{L}(\cfgnt{r}_\mathit{un}) =
  \{(\cfgt{true}\ \cfgnt{l}_\mathit{un})\}$. 
\item[Immutability:] References do not change once added to the
location map, eliminating any need for constraint rewriting as the program state evolves.
\item[Determinism:] A reference in $(\cfgnt{L}\ \cfgnt{R})$ cannot point to multiple locations simultaneously: 
$
\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\cfgnt{l} \neq \cfgnt{l}^\prime \vee \phi \neq \phi^\prime) \Rightarrow (\phi \wedge \phi^\prime = \cfgt{false}))
$
where $\cfgnt{L}^\leftarrow$ is the pre-image of $\cfgnt{L}$. 
\item[Type consistency:] All locations associated with a reference have the same type:
$\forall \cfgnt{r} \in \cfgnt{L}^\leftarrow\ (\forall (\phi\ \cfgnt{l}),(\phi^\prime\ \cfgnt{l}^\prime) \in \cfgnt{L}(r)\ (
(\mathrm{Type}\lp\cfgnt{l}\rp = \mathrm{Type}\lp\cfgnt{l}^\prime\rp)))$

@@

Other more complex rules, such as one to create a new instance of a
class, define constraints on the rewrites and more complex
transformations on the heap.
$$
\mprset{flushleft}
	\inferrule[New]{
      \cfgnt{r} = \mathrm{stack}_r\lp \rp\\
      l = \mathrm{fresh}_l\lp \cfgnt{C}\rp\\\\
      \cfgnt{R}^\prime = \cfgnt{R}[\forall \cfgnt{f} \in \mathit{fields}\lp \mathrm{C}\rp \ \lp \lp l\ \cfgnt{f}\rp  \mapsto \cfgnt{r}_\mathit{null} \rp ] \\\\
      \cfgnt{L}^\prime = \cfgnt{L}[\cfgnt{r} \mapsto \{\lp \cfgt{true}\ l\rp \}]
    }{
      \lp \cfgnt{L}\ \cfgnt{R}\ \phi\ \eta\ \lp \cfgt{new}\ \cfgnt{C}\rp \ \cfgnt{k}\rp  \rcom
      \lp \cfgnt{L}^\prime\ \cfgnt{R}^\prime\ \phi\ \eta\ \cfgnt{r}\ \cfgnt{k}\rp 
	}
$$ In this rule, when the string matches the new-expression, it is
        rewritten to use a new heap location where all of the fields
        for the new object point to $\cfgnt{r}_\mathit{null}$ and the
        location map points a new stack reference to that new object.
        Note that references for objects from the new-expression are
        auxiliary literals, so they do not appear in any constraint in
        the symbolic heap as an aliasing option. The complete set of
        semantic rules for Javalite are presented
        in Appendix A.
